Stateless Model Checking with Data-Race Preemption Points

### 2015-10-18 pldi-16/di.tex
change figures

% TODO: get latex to put this on page 2 somehow
\begin{figure*}[t]
	\includegraphics[width=0.15\textwidth]{tree0.pdf}
	\includegraphics[width=0.225\textwidth]{tree1.pdf}
	\includegraphics[width=0.25\textwidth]{tree2.pdf}
	\includegraphics[width=0.35\textwidth]{tree3.pdf}
	\caption{Iterative Deepening visualization.
		The minimal state space includes only mandatory thread switches, such as {\tt yield()} or {\tt cond\_wait()}~\cite{landslide}.
Multiple further tests can be run: preempting on calls to {\tt mutex\_lock} alone, {\tt mutex\_unlock} alone, or both together.
Each option increases the state space size by an unpredictable factor, so multiple state spaces should be tested in parallel.
Estimation techniques~\cite{estimation} inform which state spaces to prioritize.
}
	% TODO: reference this label in first few paragraphs of design
	\label{fig:id}
\end{figure*}
\section{Design}

@@

We identify a particular class of false positive data-race candidate in which the associated memory is recycled by {\tt malloc} between the two accesses.
Figure~\ref{fig:recycle} shows a common code pattern and interleaving which can expose such behavior.
If the {\tt malloc} on line 4 returns the same address passed to {\tt free} on line 2, then lines 1 and 6 will be flagged as a data race.
We term this a {\em malloc-recycle data race}.
To the human eye, this is obviously a false positive: reordering lines 4-6 before lines 1-2 will change {\tt malloc}'s return value, causing {\tt x} and {\tt y} to no longer collide.
Here, Thread 2's logic usually corresponds to the initialization pattern \cite{eraser}, but for generality we have added a {\tt publish} action on line 4.

\begin{figure}[t]
\begin{tabular}{rll}
	& \multicolumn{2}{c}{\texttt{struct x \{ int foo; int baz; \};}} \\
	& \multicolumn{2}{c}{\texttt{struct y \{ int bar; \};~~~~~~~~~~}} \\
	\\
	& Thread 1 & Thread 2 \\
	1 & \texttt{\hilight{brickred}{x->foo = ...;}} & \\
	2 & \texttt{\hilight{olivegreen}{free}(x);} \\
	3 & & \texttt{// x's memory recycled} \\
	4 & & \texttt{y~=~\hilight{olivegreen}{malloc}(...);} \\
	5 & & \texttt{publish(y);} \\
	6 & & \texttt{\hilight{brickred}{y->bar = ...;}} \\
\end{tabular}
\caption{A common execution pattern with {\tt malloc()} that produces false positive data race candidates.}
\label{fig:recycle}
\end{figure}

@@

It is quite simple to mechanically recognize when {\tt x} and {\tt y} correspond to different abstract allocations despite colliding on address.
\landslide~implements this check by adding a generation counter to its heap state tracking:
each allocation is given a unique ID, and when evaluating whether two heap accesses can race, their generation counters must match.
However, when limited to a single execution, suppressing any data race matching this pattern is unsound.
Consider the more unusual program in Figure~\ref{fig:recycle-bug}:
Now, the memory is recycled the same way, but the racing access's address is not tied to {\tt malloc}'s return value.
Hence, reordering lines 6-7 before line 3 will cause {\tt x} and {\tt x2} to race.

\begin{figure}[t]
\begin{tabular}{rll}
	& Thread 1 & Thread 2 \\
	1 & \texttt{publish(x);} & \\
	2 & & \texttt{x2 = get\_published\_x();} \\
	3 & \texttt{\hilight{brickred}{x->foo = ...;}} & \\
	4 & \texttt{\hilight{olivegreen}{free}(x);} \\
	5 & & \texttt{// x's memory recycled} \\
	6 & & \texttt{y~=~\hilight{olivegreen}{malloc}(...);} \\
	7 & & \texttt{\hilight{brickred}{x2->foo = ...;}} \\
\end{tabular}
\caption{If a single-pass data race detector discarded candidates matching the malloc-recycle pattern,
it would miss the bug in this adversarial program.}
\label{fig:recycle-bug}
\end{figure}

>>>>

% TODO: get latex to put this on page 2 somehow
\begin{figure}[t]
	\includegraphics[width=0.48\textwidth]{trees.pdf}
	%\includegraphics[width=0.15\textwidth]{tree0.pdf}
	%\includegraphics[width=0.225\textwidth]{tree1.pdf}
	%\includegraphics[width=0.25\textwidth]{tree2.pdf}
	%\includegraphics[width=0.35\textwidth]{tree3.pdf}
	\caption{Iterative Deepening example.
		The minimal state space (top) includes only mandatory thread switches, such as {\tt yield()} or {\tt cond\_wait()}~\cite{landslide}.
		Multiple further tests can be run: preempting on calls to {\tt mutex\_lock} alone (left), {\tt mutex\_unlock} alone (right), or both together (bottom).
Each option increases the state space size by an unpredictable factor, so multiple state spaces should be tested in parallel.
Estimation techniques~\cite{estimation} inform which state spaces to prioritize.
}
	% TODO: reference this label in first few paragraphs of design
	\label{fig:id}
\end{figure}
\section{Design}

@@

We identify a particular class of false positive data-race candidate in which the associated memory is recycled by {\tt malloc} between the two accesses.
Figure~\ref{fig:recycle} shows a common code pattern and interleaving which can expose such behavior.
If the {\tt malloc} on line 4 returns the same address passed to {\tt free} on line 2, then lines 1 and 7 will be flagged as a data race.
We term this a {\em malloc-recycle data race}.
To the human eye, this is obviously a false positive: reordering lines 4-7 before lines 1-2 will change {\tt malloc}'s return value, causing {\tt x} and {\tt y} to no longer collide.
Here, Thread 2's logic usually corresponds to the initialization pattern \cite{eraser}, but for generality we have added a {\tt publish} action on line 6.

\begin{figure}[t]
	\small
\begin{tabular}{rll}
	& \multicolumn{2}{c}{\texttt{struct x \{ int foo; int baz; \} *x;}} \\
	& \multicolumn{2}{c}{\texttt{struct y \{ int bar; \} *y;~~~~~~~~~~}} \\
	\\
	& Thread 1 & Thread 2 \\
	1 & \texttt{\hilight{brickred}{x->foo = ...;}} & \\
	2 & \texttt{\hilight{olivegreen}{free}(x);} \\
	3 & & \texttt{// x's memory recycled} \\
	4 & & \texttt{y~=~\hilight{olivegreen}{malloc}(sizeof *y);} \\
	5 & & \texttt{// ...initialize...}\\
	6 & & \texttt{publish(y);} \\
	7 & & \texttt{\hilight{brickred}{y->bar = ...;}} \\
\end{tabular}
\caption{A common execution pattern with {\tt malloc()} that produces false positive data race candidates.}
\label{fig:recycle}
\end{figure}

@@

It is quite simple to mechanically recognize when {\tt x} and {\tt y} correspond to different abstract allocations despite colliding on address.
\landslide~implements this check by adding a generation counter to its heap state tracking:
each allocation is given a unique ID,
and when evaluating whether two heap accesses can race,
the IDs of their containing blocks must match.
However, when limited to a single execution, suppressing any data race matching this pattern is unsound.
Consider the more unusual program in Figure~\ref{fig:recycle-bug}:
Now, the memory is recycled the same way, but the racing access's address is not tied to {\tt malloc}'s return value.
Hence, reordering lines 6-7 before line 3 will cause {\tt x} and {\tt x2} to race.

\begin{figure}[t]
	\small
\begin{tabular}{rll}
	& Thread 1 & Thread 2 \\
	1 & \texttt{publish(x);} & \\
	2 & & \texttt{x2 = get\_published\_x();} \\
	3 & \texttt{\hilight{brickred}{x->foo = ...;}} & \\
	4 & \texttt{\hilight{olivegreen}{free}(x);} \\
	5 & & \texttt{// x's memory recycled} \\
	6 & & \texttt{y~=~\hilight{olivegreen}{malloc}(sizeof *y);} \\
	7 & & \texttt{\hilight{brickred}{x2->foo = ...;}} \\
\end{tabular}
\caption{If a single-pass data race detector discarded candidates matching the malloc-recycle pattern,
it would miss the bug in this adversarial program.}
\label{fig:recycle-bug}
\end{figure}
